The role of advection versus diffusion in intraventricular transport

### 2025-03-17 paper/main.tex
Nathalie comments, update fig1

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{graphics/figure1.png}
    \caption{
    \textbf{a.} Schematic illustration of the two cerebrospinal fluid (CSF) flow components modeled,
    motile cilia and cardiac pulsations, and their contributions to the flow patterns.
    \textbf{b.} Particle tracking data visualizing flow directionality in the middle ventricle.
    \textbf{c.} Flow velocity magnitudes in the middle ventricle from particle tracking data.
    \textbf{d.} Computational mesh with marked facets indicating regions of motile cilia (magenta) and
    anterior/posterior facets (blue).
    \textbf{e.} Streamlines of the (steady-state) CSF flow field simulated with the
    cilia-only/no-cardiac flow model.
    \textbf{f.} Streamlines of the CSF flow field simulated with the cardiac-only/no-cilia flow
    model at a time 25\% into the cardiac cycle. %t=0.112613 s, period=0.45045 s
    \textbf{g.} Streamlines simulated with the baseline flow model at
    a time 25\% into the cardiac cycle. %t=0.112613 s
    \textbf{h.} Streamlines simulated with the baseline flow model at
    a time 50\% into the cardiac cycle. %t=0.225225 s
    \textbf{i.} Streamlines simulated with the baseline flow model at
    a time 75\% into the cardiac cycle. %t=0.315315 s
    \textbf{j.} Illustration of the cardiac-related normal pressure forcing imposed on the
    anterior/posterior boundaries. Time instants of panels \textbf{f, h--j} are indicated.
    }
    \label{fig:fig1}
\end{figure}

@@

Proteins, neurotransmitters and other molecular substances
intrinsically diffuse within the brain ventricles, but are also
advected by the flow of CSF. A natural question is which of these
mechanisms that dominate for any given type of solute. To address this
question, we simulate the distribution and evolution of three
different solutes injected within a region of interest (ROI1) in the
dorsal region of the middle ventricle (\Crefpanels{fig:fig2}{a}),
subject to the CSF flow induced by motile cilia and cardiac pressure
pulsations. In particular, we consider three different molecular
diffusion coefficients, corresponding to a large exosome ($D_1 = 1.63
\times 10^{-12}$ m$^2$/s), a 90 kDa protein STM-GFP ($D_2 = 5.75
\times 10^{-11}$ m$^2$/s), and a 26 kDa photoconvertible protein
Dendra2 ($D_3 = 1.15 \times 10^{-10}$ m$^2$/s).

@@

Interestingly, the distribution patterns differ between the
solutes. For the smallest diffusion coefficient, representing
exosomes, the advective transport by the CSF dominates diffusion: the
CSF flow structures carry over into the concentration field,
indicating that transport along the streamlines is more rapid than the
diffusion across the streamlines (\Crefpanels{fig:fig2}{b}). The
balance between diffusion and advection shifts with increasing
diffusion coefficient as the solute spreads more uniformly throughout
the ventricular geometry (\Crefpanels{fig:fig2}{b--d}). Notably, the
smaller solutes with higher diffusion coefficients spread more quickly
to distal regions. Computing Péclet numbers, we find that
$\mathrm{Pe}_1=519$, $\mathrm{Pe}_2=14.7$ and $\mathrm{Pe}_3=7.4$ for
$D_1$, $D_2$ and $D_3$, respectively. With all $\mathrm{Pe} > 1$, this
indicates that advection is the dominant transport mechanism on a
global scale.

@@

\begin{figure}%[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure2_simulation_results_different_coeffs.png}\caption{Simulated distribution and evolution after photoconversion of
    exosomes ($D_1$), STM-GFP ($D_2$) and Dendra2 ($D_3$). 
    \textbf{a.} Geometry-centered clip of the ventricles mesh showing the locations of the ROIs.
    \textbf{b--d.} Slice of the concentration field simulated
    with diffusion coefficients $D_1$, $D_2$ and $D_3$ for eight time instants covering the
    time range from 45 seconds to 790 seconds, corresponding to 100 to 1750 cardiac cycles.
    Slices are taken in the $xz$-plane ($y=0.140$ mm, the center of the geometry).
    The colorbar is scaled so the maximum is the maximum concentration of the Dendra2 
    concentration field slice at the last time instant.
    \textbf{e--g.} Mean concentration $\cbar$ in each region of interest (ROI)
    as function of time for diffusion coefficient $D_1$, $D_2$ and $D_3$, respectively. 
    \textbf{h.} The time $\hat{t}$ before the mean concentration $\cbar$ exceeds a
    threshold value of $0.25$ (ROIs 2-4) or $0.10$ (ROIs 5, 6) in each simulation setup.
    Note that there is no bar for $D_1$ in ROI 6, because the threshold value was never exceeded.}\label{fig:fig2}
\end{figure}

To further quantify the solute transport, we examine the mean
concentrations $\cbar_i(t)$ in the six ROIs over time
(\Crefpanels{fig:fig2}{e--g}) and, in addition, the times $\hat{t}_i$
when the mean concentrations first exceed a threshold value
$\hat{c}_i$ (\Crefpanels{fig:fig2}{h}).  Higher diffusion coefficients
result in more rapid transport (\Crefpanels{fig:fig2}{e--g}), with
lower times-to-threshold in ROIs 2--6 (\Crefpanels{fig:fig2}{h}). For
all three diffusion coefficients, the solute first spreads within the
middle ventricle, covering ROIs 2--4, before appearing in the anterior
ventricle (ROI5) and finally reaching the posterior ventricle
(ROI6). Notably, the exosomes, associated with the smallest diffusion
coefficient, do not reach the posterior ventricle (ROI6) within the
simulated time frame.

@@

To validate the model predictions, we performed photoconversion
experiments in transgenic zebrafish larvae expressing the secreted
Dendra2 in their brain ventricles. Our photoconversion protocol
consisted of the acquisition of a baseline fluorescence intensity
$F_0$, before localized exposure to a UV laser (at ROI 1) that
converts Dendra2 from a green- to a red-emitting fluorescent
protein. To quantify the transport of photoconverted proteins, we
measured and calculated the change in fluorescence intensity $\Delta F
= (F(t)-F_0)/F_0$ over time (\Crefpanels{fig:fig3}{a}).

@@

\subsection{Absence of ciliary motion affects local protein distribution}

@@

We previously identified three distinct ciliated cell lineages in the
zebrafish larval brain~\cite{Olstad2019CiliaryDevelopment, DGama2025MotileBrain}.
To identify the effects of these lineages on CSF flow and transport, 
we investigate region-specific cilia paralysis with the computational model.
Three cilia-covered regions were considered paralyzed separately:
the telencephalic (anterior) ventricle, and the dorsal and ventral parts
of the diencephalic (middle) ventricle (\Crefpanels{fig:fig5}{a}).
To simulate cilia paralysis, the cilia tangential traction boundary condition
is for each case changed to a free-slip boundary condition
in the paralyzed region,
prescribing zero tangential forces to
represent absence of cilia motion.

@@

\subsection{Modification of ventricles geometry impacts protein distribution}
Variations in ventricular morphology impact CSF flow patterns.
To assess how these variations influence solute distribution,
we consider four alternative brain ventricles meshes:
one where the duct connecting the anterior and middle ventricles is constricted
(66 \% reduction in cross-sectional area);
one where the duct connecting the middle and posterior ventricles is constricted
(33 \% reduction);
one where the middle ventricle is shrunk in the lateral direction (43 \% reduction);
and the combination of the above three modifications (\Crefpanels{fig:fig6}{a, b}). In these
four modified geometries the volume is reduced by 3.9\%, 4.1\%, 1.4\% and 9.4\%, respectively,
compared to the original geometry.

@@

For the mesh where all three ventricles are modified,
we observe higher retention of solute in the middle ventricle (ROIs 2--4),
while both the anterior and the posterior ventricles (ROIs 5 and 6) have reduced
mean concentrations. Altogether, our results show that the geometry and
width of the ventricular ducts matter
for solute transport, with notable differences in transport from relatively small
geometry modifications. This may explain why we observed larger differences
in the cilia-deficient computational model compared to the experimental data in
\Cref{fig:fig4}, as the model and the fish have different duct diameters.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure6_compare_modified_geometries.png}
    \caption{\textbf{a.} The original computational mesh of the ventricular geometry.
    \textbf{b.} The four modified versions of the computational geometry. Compared
    to the original mesh, the volumes of the deformed meshes are reduced by 3.9\%, 4.1\%, 1.4\% 
    and 9.4\%.
    \textbf{c.} Mean concentration $\cbar(t)$ simulated with the original and modified geometries
    in the 6 ROIs.
    \textbf{d.} A comparison of the tangential cilia force in the three cilia regions
    and the total force applied, calculated with \Cref{eq:tangential_force_calculation}.
    In each region, the force $F_{\mathrm{original}}$ 
    applied on the original mesh is the baseline of 100\%,
    to which the forces $F_i$ of the other mesh versions are compared, where $i$ denotes
    the cases fore-shrunk, middle shrunk, mid-hind shrunk, and the combination of the three.
    The bars for the four shrunk scenarios thus
    display $F_i/F_{\mathrm{original}}\cdot 100\%$.}
    \label{fig:fig6}
\end{figure}

\section{Discussion}
In this study, we have presented a computational model of CSF flow and solute transport within the brain ventricles.
We used the larval zebrafish brain ventricles since its geometry,
flow components and kinetics are well described, thereby allowing us to validate
our simulation experimentally. Nevertheless, our computational framework can be adapted to and implemented in
other ventricular systems in the future, given that the flow parameters of such systems
are comprehensively resolved. By imposing tangential traction and normal pressure
on the CSF flow, we were able to reproduce flow fields with similar features
as observed experimentally. The prominent flow features were vortex-structured flow,
mediated by cilia, and pulsatile flow, mediated by the cardiac cycle.

@@

\subsection{The roles of advection and diffusion in intraventricular solute transport}
To investigate the roles of advection and diffusion in solute transport, we quantified transport
based on changes in fluorescence intensity $\Delta F$ (experiments) and
mean concentration $\cbar$ (simulations) as functions of time.
Even though we identified very similar dynamics in simulations and
in vivo experiments, the dynamics of the simulations were generally
more rapid and resulted in higher total amounts of solute.
Differences between experiments and simulations may relate to
ventricular size, as we showed that modifying the geometry of the interventricular ducts
impacts solute distribution, and the geometry used for the computational model was larger
than geometries used in experiments. Alternatively, dissimilarities in results may relate
to the way we defined the photoconversion site (ROI 1) in the simulations,
since the volume of the photoconversion region determines
the total amount of solute and thereby the rate of transport. Another source
of discrepancy may relate to the permeability of the ventricular wall,
which we have ignored in our simulation. Reports have shown that the
neuroepithelium lining the walls of brain ventricles in 24 hour old zebrafish
is permeable to dyes smaller than 70 kDa~\cite{Chang2012AnNeuroepithelium}.
Since Dendra2 is around 26 kDa in size~\cite{Gurskaya2006EngineeringLight} and
thus under this reported limit, one could assume that Dendra2 may
leak out of the brain ventricles over the time spans we consider
and render the dynamics slower.

@@

Meanwhile, we identified that transport of larger molecules
(thus with smaller diffusion coefficients),
such as 150 nm exosomes, are almost entirely dependent on advection.
Indeed, there was essentially zero solute in the posterior ventricle at the
final simulation time (around 860 s), a ventricle where transport is largely driven
by diffusion due to the absence of motile cilia.
For $D_1$, an estimate of the diffusion timescale $t_d=L^2/D_1$ over a distance
$L=200$ \textmu m from the middle to the posterior ventricle is approximately 24500 s.
Advective transport by the mean velocity $\overline{U}=2.4 \ \mu$m/s over the same
distance would happen on a timescale of the order of $t_a=L/\overline{U}=83$ s.
Comparing this with the minuscule values of $\cbar_6$ after simulating transport
with $D_1$ for 850 s, one can conclude that if advection
played a role in transporting solute from the photoconversion site in the middle ventricle
to the posterior ventricle, values of $\cbar_6$ would be higher than those observed.

\subsection{Effects of cilia motility and ventricular morphology}
To gain further insights into how CSF flow and intraventricular transport depend on cilia
motility and ventricular morphology, we manipulated certain
parameters in our models. First, we investigated how complete loss of cilia motility
impacted the transport dynamics. Simulations with a cardiac-only/no-cilia lead to
slower transport when compared to the baseline model that included cilia. Especially
the middle-ventricular region downstream of CSF flow (ROI 2) and the anterior ventricle (ROI 5)
experienced substantially increased times-to-threshold. We also observed
slower dynamics in the rest of the middle ventricle (ROIs 3 and 4), whereas the
posterior ventricle (ROI 6) was unaffected by the loss
of cilia motility, indicating that transport from the middle to the posterior ventricle is
purely driven by diffusion. We also performed photoconversion experiments on cilia mutant
zebrafish larva (\emph{smh}), observing statistically significant slower transport in 
ROIs 2 and 4. The results provide not only confidence in that our computational model
well represents the physics, but also valuable quantification of the impact of cilia motility
on intraventricular transport. Interestingly, a previous simulation study of reduced cilia motility
also observed a drastic reduction in CSF mixing in specific regions of the brain
ventricles~\cite{Yoshida2022EffectVentricles}, similar to our results. Since impaired brain 
clearance of tracers has been demonstrated in idiopathic normal pressure hydrocephalus
patients~\cite{Eide2020MagneticHydrocephalus}, cilia-enhanced mixing in the brain might be
fundamental to physiological transport within the brain.

@@

To simulate protein transport in the brain ventricles, we numerically solved
an advection-diffusion equation. Except for where we imposed a normal pressure
boundary condition on the CSF flow,
we imposed a no-flux condition on the concentration. Imposing a pressure-driven
cardiac-related flow on certain parts of the CSF flow boundary leads to in- and outflow,
and the advective transport then requires handling of in- and outflux boundary conditions.
We set an approximate periodic boundary condition on the the in- and outflux boundaries,
such that the total mass is conserved in the
limit that the timestep size is zero, reflecting our assumption that the
ventricular system is closed.
We handled the outflux by leaving the advective flux term as an unknown in the weak formulation
of the advection-diffusion equation. This has been referred to with different names
in previous works, names such as the ``free boundary condition"~\cite{Papanastasiou1992ACondition},
the ``no boundary condition"~\cite{Griffiths1997TheCondition}, and the 
``consistent flux boundary condition"~\cite{Lynch2020NumericalHemodynamics}. 
In leaving the advective flux term in the weak form we are lacking a boundary condition,
which is problematic in the continuous sense of the problem.
However, the discrete problem is well-posed, because the approach represents
constraining higher-order derivatives of the concentration~\cite{Griffiths1997TheCondition}. 
We deliberately excluded the space proximal to in- and outflow boundaries in
the anterior and posterior regions of interest (5 and 6) to limit adverse effects of 
the advective transport at the in- and outflux boundaries on the mean concentrations.

A possible improvement to the in- and outflow boundary conditions we employed would be
to consider a 3D-0D coupling of the in- and outflow boundaries using
ordinary differential equations,
similar to what has been done in cardiac
modeling~\cite{Brown2024AMechanics, Augustin2021ACirculation}.
However, if the ventricular system in reality is a closed system,
one should rather seek out a way of modeling the CSF flow without open boundaries to avoid
handling in- and outflow boundary conditions. A more appropriate way of handling the beating
cardiac motion could be to impose a ventricular wall deformation, assumed to result from pulsatile
motion of the heart and/or cardiac vessels surrounding the ventricles. This wall deformation would
induce CSF flow while simultaneously permitting enforcement of the impermeability condition
$\uu\cdot\nn=0$ on the ventricular walls. Such wall-motion induced CSF flow has been modeled and
simulated successfully in the human brain ventricular system~\cite{Causemann2022HumanFramework,
Kurtcuoglu2005ComputationalSystem, Kurtcuoglu2007ComputationalSylvius, Linninger2005PulsatileBrain}.

@@

In large brains such as the human brain, we presume ciliary beating mostly 
contributes to flow in localized regions proximal to the ventricular walls due to
the microscopic length of the cilia.
Although the precise mechanisms of intraventricular flow are still debated,
much evidence indicate that blood vessel pulsations, respiration, CSF secretion
and brain motion induce the pulsatile bulk flow in the central
regions of the ventricles~\cite{Vinje2019RespiratoryMeasurements,
Siyahhan2014FlowVentricles, Linninger2005PulsatileBrain, Enzmann1992BrainImaging,
Kurtcuoglu2007ComputationalSylvius}.
In human brains, cilia therefore probably play an important role in regulating the
amount of substances like exosomes close to the ventricular walls, similar to coral
reefs, where it has been observed that cilia are primary regulators of
oxygen amounts at the coral surface~\cite{Pacherres2022CiliaryProduction}.
Flow mediated by cilia indeed serves such a purpose in nodal flow at early
developmental stages of the brain~\cite{Hirokawa2006NodalAsymmetry}, and the discovery
of a cilia-driven flow network in rodent brains suggest such selective
transport~\cite{Faubel2016Cilia-basedVentricles}.

>>>>

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{graphics/figure1.png}
    \caption{
    \textbf{a.} Schematic illustration of the two cerebrospinal fluid (CSF) flow components modeled,
    motile cilia and cardiac pulsations, and their contributions to the flow patterns.
    \textbf{b, c.} Particle tracking data visualizing CSF flow directionality (\textbf{b}) and
    velocity magnitudes in the middle ventricle.
    \textbf{d.} Computational mesh with marked facets indicating regions of motile cilia (magenta) and
    anterior/posterior facets (blue).
    \textbf{e.} Streamlines of the (steady-state) CSF flow field simulated with the
    cilia-only/no-cardiac flow model.
    \textbf{f.} Streamlines of the CSF flow field simulated with the cardiac-only/no-cilia flow
    model at a time 25\% into the cardiac cycle. %t=0.112613 s, period=0.45045 s
    \textbf{g.} Illustration of the cardiac-related normal pressure forcing imposed on the
    anterior/posterior boundaries.
    \textbf{h--j.} Streamlines simulated with the baseline flow model at
    a time a time 25\% (\textbf{h}), 50\% (\textbf{i}) and 75\% (\texbf{j})
    into the cardiac cycle. Times indicated in panel \textbf{g}.
    %t=0.112613 s %t=0.225225 s %t=0.315315 s
    Scale bars 50 \textmu m.
    }
    \label{fig:fig1}
\end{figure}

@@

Proteins, neurotransmitters and other molecular substances
intrinsically diffuse within the brain ventricles, but are also
advected by the flow of CSF. A natural question is which of these
mechanisms that dominate for any given type of solute. To address this
question, we simulate the distribution and evolution of three
different solutes injected within a region of interest (ROI1) in the
dorsal region of the middle ventricle (\Crefpanels{fig:fig2}{a}),
subject to the CSF flow induced by motile cilia and cardiac pressure
pulsations. In particular, we consider three different molecular
diffusion coefficients, corresponding to a large extracellular vesicle ($D_1 = 2.17
\times 10^{-12}$ m$^2$/s), a 90 kDa protein STM-GFP ($D_2 = 5.75
\times 10^{-11}$ m$^2$/s), and a 26 kDa photoconvertible protein
Dendra2 ($D_3 = 1.15 \times 10^{-10}$ m$^2$/s).

@@

Interestingly, the distribution patterns differ between the
solutes. For the smallest diffusion coefficient, representing
extracellular vesicles, the advective transport by the CSF dominates diffusion: the
CSF flow structures carry over into the concentration field,
indicating that transport along the streamlines is more rapid than the
diffusion across the streamlines (\Crefpanels{fig:fig2}{b}). The
balance between diffusion and advection shifts with increasing
diffusion coefficient as the solute spreads more uniformly throughout
the ventricular geometry (\Crefpanels{fig:fig2}{b--d}). Notably, the
smaller solutes with higher diffusion coefficients spread more quickly
to distal regions. Computing Péclet numbers, we find that
$\mathrm{Pe}_1=519$, $\mathrm{Pe}_2=14.7$ and $\mathrm{Pe}_3=7.4$ for
$D_1$, $D_2$ and $D_3$, respectively. With all $\mathrm{Pe} > 1$, this
indicates that advection is the dominant transport mechanism on a
global scale.

@@

\begin{figure}%[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure2_simulation_results_different_coeffs.png}\caption{Simulated distribution and evolution after photoconversion of
    extracellular vesicles (EVs) ($D_1$), STM-GFP ($D_2$) and Dendra2 ($D_3$). 
    \textbf{a.} Geometry-centered clip of the ventricles mesh showing the locations of the ROIs.
    \textbf{b--d.} Slice of the concentration field simulated
    with diffusion coefficients $D_1$, $D_2$ and $D_3$ for eight time instants covering the
    time range from 45 seconds to 790 seconds, corresponding to 100 to 1750 cardiac cycles.
    Slices are taken in the $xz$-plane ($y=0.140$ mm, the center of the geometry).
    The colorbar is scaled so the maximum is the maximum concentration of the Dendra2 
    concentration field slice at the last time instant.
    \textbf{e--g.} Mean concentration $\cbar$ in each region of interest (ROI)
    as function of time for diffusion coefficient $D_1$, $D_2$ and $D_3$, respectively. 
    \textbf{h.} The time $\hat{t}$ before the mean concentration $\cbar$ exceeds a
    threshold value of $0.25$ (ROIs 2-4) or $0.10$ (ROIs 5, 6) in each simulation setup.
    Note that there is no bar for $D_1$ in ROI 6, because the threshold value was never exceeded.}\label{fig:fig2}
\end{figure}

To further quantify the solute transport, we examine the mean
concentrations $\cbar_i(t)$ in the six ROIs over time
(\Crefpanels{fig:fig2}{e--g}) and, in addition, the times $\hat{t}_i$
when the mean concentrations first exceed a threshold value
$\hat{c}_i$ (\Crefpanels{fig:fig2}{h}).  Higher diffusion coefficients
result in more rapid transport (\Crefpanels{fig:fig2}{e--g}), with
lower times-to-threshold in ROIs 2--6 (\Crefpanels{fig:fig2}{h}). For
all three diffusion coefficients, the solute first spreads within the
middle ventricle, covering ROIs 2--4, before appearing in the anterior
ventricle (ROI5) and finally reaching the posterior ventricle
(ROI6). Notably, the extracellular vesicles, associated with the smallest diffusion
coefficient, do not reach the posterior ventricle (ROI6) within the
simulated time frame.

@@

To validate the model predictions, we performed photoconversion
experiments in transgenic zebrafish larvae expressing the secreted
Dendra2 in their brain ventricles. The approach is inspired by earlier work
where a photoconvertible protein was injected into the ventricle~\cite{fame2016directional}.
By using a transgenic system where the protein is directly secreted into the CSF,
we avoid detrimental effects of intraventricular injection on CSF properties. 
Our photoconversion protocol
consisted of the acquisition of a baseline fluorescence intensity
$F_0$, before localized exposure to a UV laser (at ROI 1) that
converts Dendra2 from a green- to a red-emitting fluorescent
protein. To quantify the transport of photoconverted proteins, we
measured and calculated the change in fluorescence intensity $\Delta F
= (F(t)-F_0)/F_0$ over time (\Crefpanels{fig:fig3}{a}).

@@

\subsection{Absence of ciliary motion affects local solute distribution}

@@

We previously characterized three distinct ciliated cell lineages in the
zebrafish larval brain~\cite{Olstad2019CiliaryDevelopment, DGama2025MotileBrain}.
To identify the effects of these lineages on CSF flow and transport, 
we investigate region-specific cilia paralysis with the computational model.
Three cilia-covered regions were considered paralyzed separately:
the telencephalic (anterior) ventricle, and the dorsal and ventral parts
of the diencephalic (middle) ventricle (\Crefpanels{fig:fig5}{a}).
To simulate cilia paralysis, the cilia tangential traction boundary condition
is for each case changed to a free-slip boundary condition
in the paralyzed region,
prescribing zero tangential forces to
represent absence of cilia motion.

@@

\subsection{Modification of ventricular geometry impacts solute distribution}
Variations in ventricular morphology impact CSF flow patterns.
To assess how these variations influence solute distribution,
we consider four alternative brain ventricles meshes:
one where the duct connecting the anterior and middle ventricles is constricted
(66 \% reduction in cross-sectional area);
one where the duct connecting the middle and posterior ventricles is constricted
(33 \% reduction);
one where the middle ventricle is shrunk in the lateral direction (43 \% reduction);
and the combination of the above three modifications (\Crefpanels{fig:fig6}{a, b}). In these
four modified geometries the volume is reduced by 3.9\%, 4.1\%, 1.4\% and 9.4\%, respectively,
compared to the original geometry.

@@

For the mesh where all three ventricles are modified,
we observe higher retention of solute in the middle ventricle (ROIs 2--4),
while both the anterior and the posterior ventricles (ROIs 5 and 6) have reduced
mean concentrations. Altogether, our results show that the geometry and
width of the ventricular ducts matter
for solute transport, with notable differences in transport from relatively small
geometry modifications. This may explain why we observed larger differences
in the cilia-deficient computational model compared to the experimental data in
\Cref{fig:fig4}, as the model and the fish have slightly different duct diameters.
\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{graphics/figure6_compare_modified_geometries.png}
    \caption{\textbf{a.} The original computational mesh of the ventricular geometry.
    \textbf{b.} The four modified versions of the computational geometry. Compared
    to the original mesh, the volumes of the deformed meshes are reduced by 3.9\%, 4.1\%, 1.4\% 
    and 9.4\%.
    \textbf{c.} Mean concentration $\cbar(t)$ simulated with the original and modified geometries
    in the 6 ROIs.
    \textbf{d.} A comparison of the tangential cilia force in the three cilia regions
    and the total force applied, calculated with \Cref{eq:tangential_force_calculation}.
    In each region, the force $F_{\mathrm{original}}$ 
    applied on the original mesh is the baseline of 100\%,
    to which the forces $F_i$ of the other mesh versions are compared, where $i$ denotes
    the cases fore-shrunk, middle shrunk, mid-hind shrunk, and the combination of the three.
    The bars for the four shrunk scenarios thus
    display $F_i/F_{\mathrm{original}}\cdot 100\%$.}
    \label{fig:fig6}
\end{figure}

\section{Discussion}
In this study, we have presented a computational model of CSF flow and solute transport within the brain ventricles.
We used the larval zebrafish brain ventricles since its geometry,
flow components and kinetics are well described, thereby allowing us to validate
our simulation experimentally. Nevertheless, our computational framework can be adapted to and implemented in
other ventricular systems in the future, once the flow parameters of such systems
are comprehensively resolved. By imposing tangential traction and normal pressure
on the CSF flow, we were able to reproduce flow fields with similar features
as observed experimentally. The prominent flow features were vortex-structured flow,
mediated by cilia, and pulsatile flow, mediated by the cardiac cycle.

@@

\subsection{The roles of advection and diffusion in intraventricular solute transport}
To investigate the roles of advection and diffusion in solute transport, we quantified transport
based on changes in fluorescence intensity $\Delta F$ (experiments) and
mean concentration $\cbar$ (simulations) as functions of time.
Even though we identified very similar dynamics in simulations and
in vivo experiments, the dynamics of the simulations were generally
more rapid and resulted in higher total amounts of solute.
Differences between experiments and simulations may relate to
ventricular size, as we showed that modifying the geometry of the interventricular ducts
impacts solute distribution, and the geometry used for the computational model was larger
than some of the geometries used in experiments.
Notably, there are large variations in ventricular size in the control population at 2 dpf.
Alternatively, dissimilarities in results may relate
to the way we defined the photoconversion site (ROI 1) in the simulations,
since the volume of the photoconversion region determines
the total amount of solute and thereby the rate of transport. Another source
of discrepancy may relate to the permeability of the ventricular wall,
which we have ignored in our simulation. Reports have shown that the
neuroepithelium lining the walls of brain ventricles in 24 hour old zebrafish
is permeable to dyes smaller than 70 kDa~\cite{Chang2012AnNeuroepithelium}.
Since Dendra2 is around 26 kDa in size~\cite{Gurskaya2006EngineeringLight} and
thus under this reported limit, one could assume that Dendra2 may
leak out of the brain ventricles over the time spans we consider
and render the dynamics slower.

@@

Meanwhile, we identified that transport of larger molecules
(thus with smaller diffusion coefficients),
such as 150 nm radius extracellular vesicles, are almost entirely dependent on advection.
Indeed, there was essentially zero solute in the posterior ventricle at the
final simulation time (around 860 s), a ventricle where transport is largely driven
by diffusion due to the absence of motile cilia.
For $D_1$, an estimate of the diffusion timescale $t_d=L^2/D_1$ over a distance
$L=200$ \textmu m from the middle to the posterior ventricle is approximately 24500 s.
Advective transport by the mean velocity $\overline{U}=2.4 \ \mu$m/s over the same
distance would happen on a timescale of the order of $t_a=L/\overline{U}=83$ s.
Comparing this with the minuscule values of $\cbar_6$ after simulating transport
with $D_1$ for 850 s, one can conclude that if advection
played a role in transporting solute from the photoconversion site in the middle ventricle
to the posterior ventricle, values of $\cbar_6$ would be higher than those observed.

\subsection{Effects of cilia motility and ventricular morphology}
To gain further insights into how CSF flow and intraventricular transport depend on cilia
motility and ventricular morphology, we manipulated certain
parameters in our models. First, we investigated how complete loss of cilia motility
impacted the transport dynamics. Simulations with a cardiac-only/no-cilia lead to
slower transport when compared to the baseline model that included cilia. Especially
the middle-ventricular region downstream of CSF flow (ROI 2) and the anterior ventricle (ROI 5)
experienced substantially increased times-to-threshold. We also observed
slower dynamics in the rest of the middle ventricle (ROIs 3 and 4), whereas the
posterior ventricle (ROI 6) was unaffected by the loss
of cilia motility, indicating that transport from the middle to the posterior ventricle is
purely driven by diffusion at this developmental stage.
We also performed photoconversion experiments on cilia mutant
zebrafish larva (\emph{smh}), observing statistically significant slower transport in 
ROIs 2 and 4. The results provide not only confidence in that our computational model
represents the physics well, but also valuable quantification of the impact of cilia motility
on intraventricular transport. Interestingly, a previous simulation study of reduced cilia motility
also observed a drastic reduction in CSF mixing in specific regions of the brain
ventricles~\cite{Yoshida2022EffectVentricles}, similar to our results. Since impaired brain 
clearance of tracers has been demonstrated in idiopathic normal pressure hydrocephalus
patients~\cite{Eide2020MagneticHydrocephalus}, cilia-enhanced mixing in the brain might be
fundamental to physiological transport within the brain.

@@

To simulate protein transport in the brain ventricles, we numerically solved
an advection-diffusion equation. Except for where we imposed a normal pressure
boundary condition on the CSF flow,
we imposed a no-flux condition on the concentration. Imposing a pressure-driven
cardiac-related flow on certain parts of the CSF flow boundary leads to in- and outflow,
and the advective transport then requires handling of in- and outflux boundary conditions.
We set an approximate periodic boundary condition on the the in- and outflux boundaries,
such that the total mass is conserved in the
limit that the timestep size is zero, reflecting our assumption that the
ventricular system is closed.
We handled the outflux by leaving the advective flux term as an unknown in the weak formulation
of the advection-diffusion equation. This has been referred to with different names
in previous works, names such as the ``free boundary condition"~\cite{Papanastasiou1992ACondition},
the ``no boundary condition"~\cite{Griffiths1997TheCondition}, and the 
``consistent flux boundary condition"~\cite{Lynch2020NumericalHemodynamics}. 
In leaving the advective flux term in the weak form we are lacking a boundary condition,
which is problematic in the continuous sense of the problem.
However, the discrete problem is well-posed, because the approach represents
constraining higher-order derivatives of the concentration~\cite{Griffiths1997TheCondition}. 
We deliberately excluded the space proximal to in- and outflow boundaries in
the anterior and posterior ventricles (ROIs 5 and 6) to limit adverse effects of 
the advective transport at the in- and outflux boundaries on the mean concentrations.

A possible improvement to the in- and outflow boundary conditions would be
to consider a 3D-0D coupling of the in- and outflow boundaries using
ordinary differential equations,
similar to what has been done in cardiac
modeling~\cite{Brown2024AMechanics, Augustin2021ACirculation}.
However, if the ventricular system in reality is a closed system,
one should rather seek out a way of modeling the CSF flow without open boundaries to avoid
handling in- and outflow boundary conditions. A more appropriate way of handling the beating
cardiac motion could be to impose a ventricular wall deformation, assumed to result from pulsatile
motion of the heart and/or cardiac vessels surrounding the ventricles. This wall deformation would
induce CSF flow while simultaneously permitting enforcement of the impermeability condition
$\uu\cdot\nn=0$ on the ventricular walls. Such wall-motion induced CSF flow has been modeled and
simulated successfully in the human brain ventricular system~\cite{Causemann2022HumanFramework,
Kurtcuoglu2005ComputationalSystem, Kurtcuoglu2007ComputationalSylvius, Linninger2005PulsatileBrain}.

@@

In large brains such as the human brain, we presume ciliary beating mostly 
contributes to flow in localized regions proximal to the ventricular walls due to
the microscopic length of the cilia.
Although the precise mechanisms of intraventricular flow are still debated,
much evidence indicate that blood vessel pulsations, respiration, CSF secretion
and brain motion induce the pulsatile bulk flow in the central
regions of the ventricles~\cite{Vinje2019RespiratoryMeasurements,
Siyahhan2014FlowVentricles, Linninger2005PulsatileBrain, Enzmann1992BrainImaging,
Kurtcuoglu2007ComputationalSylvius}.
In human brains, cilia therefore probably play an important role in regulating the
amount of substances like extracellular vesicles close to the ventricular walls, similar to coral
reefs, where it has been observed that cilia are primary regulators of
oxygen amounts at the coral surface~\cite{Pacherres2022CiliaryProduction}.
Flow mediated by cilia indeed serves such a purpose in nodal flow at early
developmental stages of the brain~\cite{Hirokawa2006NodalAsymmetry}, and the discovery
of a cilia-driven flow network in rodent brains suggest such selective
transport~\cite{Faubel2016Cilia-basedVentricles}.
